\chapter{Introduction}

Breast cancer continues to be one of the most pressing challenges in modern
healthcare, representing the most frequently diagnosed cancer among women
worldwide.
In 2022 alone, approximately 2.3 million new cases were recorded, alongside
666,000 deaths, illustrating its substantial impact as a global health issue.
Accounting for 23.8\% of all cancer diagnoses in
women\supercite{bray_global_2024,ferlay_global_2024}, breast cancer remains the
leading cause of cancer-related death for women.
Projections suggest an alarming rise in incidence, with estimates reaching 4.4
million cases annually by 2070, placing even greater strain on healthcare
systems\supercite{lei_global_2021}.

Despite advancements in screening, early detection, and treatment, disparities
in access to care persist, particularly in low-resource settings.
These disparities contribute to late-stage diagnoses and higher mortality rates
in such regions\supercite{wilkinson_understanding_2022,ginsburg_breast_2020}.
Globally, the age-standardized mortality rate for breast cancer remains high at
approximately 12.65 deaths per 100,000 women per year, highlighting the ongoing
challenges in managing this complex
disease\supercite{bray_global_2024,ferlay_global_2024}.
Given the persistence of these challenges, significant efforts are required in
research, public health strategies, and healthcare delivery to improve patient
outcomes\supercite{desantis_breast_2019}.

At the molecular level, an emerging area of research involves circular RNAs
(circRNAs), a novel class of RNA molecules characterized by their covalently
closed loop structures.
Unlike linear RNAs, circRNAs exhibit increased stability and resistance to
degradation, making them reliable biomarkers and potential therapeutic targets
across a variety of diseases, including
cancer\supercite{ma_circular_2020,hoque_exploring_2023,wilusz_circular_2017}.
Initially considered byproducts of mRNA splicing, circRNAs are now recognized
for their regulatory roles in gene expression, cellular processes, and disease
mechanisms\supercite{cherubini_foxp1_2019,wilusz_360_2018}.

The unique characteristics of circRNAs — stability, abundance, and
tissue-specific expression — have made them attractive candidates for
diagnostic and therapeutic applications, particularly in cancer.
CircRNAs show promise as biomarkers for tumor progression and treatment
response\supercite{bao_prognostic_2020,ren_construction_2017}, and their
stability in bodily fluids enhances their utility in non-invasive liquid
biopsies\supercite{bao_prognostic_2020,zhang_circular_2018}.
In breast cancer specifically, circRNAs have been identified as key players in
disease progression, particularly in relation to estrogen receptor (ER)
signaling.
Dysregulated circRNA expression has been observed in breast cancer, with
circRNAs functioning as competing endogenous RNAs (ceRNAs) that sponge miRNAs
involved in crucial pathways like ER
signaling\supercite{nair_circular_2016,xu_circrna_2022}.
Certain circRNAs, such as circRNA-SFMBT2, have been implicated in tamoxifen
resistance, a major hurdle in the treatment of ER-positive breast
cancer\supercite{li_circrna-sfmbt2_2023}.
Additionally, circRNAs like circTADA2As and circFBXL5 have been shown to
regulate miRNAs involved in key signaling pathways that promote cell
proliferation and metastasis, further highlighting their clinical
significance\supercite{xu_circtada2as_2019,gao_hsa_circrna_0006528_2019}.

This thesis seeks to further explore the role of circRNAs in breast cancer,
with a specific focus on their involvement in estrogen signaling and tumor
progression.
To achieve this, sequencing data from two mouse model studies conducted by
Furth et al.
\supercite{furth_esr1_2023,furth_overexpression_2023}
will be analyzed, examining circRNA expression profiles in relation to
variables such as age, transgene induction, and anti-hormone treatments.

To identify and quantify circRNAs from the RNA-seq data, the nf-core/circrna
pipeline\supercite{digby_nf-corecircrna_2023} will be employed, followed by
differential expression analysis to uncover circRNAs that are differentially
expressed across various experimental conditions.
Additionally, this thesis will compare different tools and methodologies for
circRNA identification, quantification, and differential expression analysis,
assessing their performance and reliability in capturing circRNA dynamics.

The structure of this thesis is designed to guide the reader through the key
concepts, methodologies, and findings related to the role of circRNAs in breast
cancer.
The Background section (\cref{chap:background}) provides the foundational
knowledge necessary to understand the subsequent chapters on methodology and
analysis.
It begins with an overview of breast cancer epidemiology (\cref{sec:brca}),
types (\cref{sec:brca_types}), causes and risk factors
(\cref{sec:brca_risk-factors}), diagnosis (\cref{sec:brca_diagnosis}), and
treatment (\cref{sec:brca_treatment}), followed by an introduction to estrogen
signaling (\cref{sec:estrogen_signaling}), which is central to breast cancer
biology.
Next, the basic mechanisms of gene expression are explained
(\cref{sec:gene_expression}), laying the groundwork for understanding the
complex roles of RNA molecules.
This is followed by a dedicated overview of circular RNAs
(\cref{sec:circrnas}), covering their biogenesis
(\cref{sec:circrna_biogenesis}), types (\cref{sec:circrna_types}), functions
(\cref{sec:circrna_functions}), and potential applications
(\cref{sec:circrna_applications}) in both molecular biology and cancer
research.

In the Materials and Methods section (\cref{chap:materials_and_methods}), the
thesis details the mouse model studies by Furth et al.
\supercite{furth_esr1_2023,furth_overexpression_2023} (\cref{sec:data}),
which form the basis for the circRNA expression
analysis.
Additionally, this section describes the nf-core/circrna
pipeline\supercite{digby_nf-corecircrna_2023} used to process and analyze the
RNA-seq data (\cref{sec:nf-core_circrna}).

The Results and Discussion section then presents a comparison of different
tools used for circRNA identification, quantification, and differential
expression analysis.
This is followed by an in-depth analysis of circRNA expression profiles
observed in the mouse models, with a focus on the implications for breast
cancer progression and estrogen signaling.
