\section{Data}
This section describes the data used in the study.

\subsection{Mouse models}
Describe the mouse models used in the study and how they were generated.

\subsection{Samples}
Describe the samples used in the study and how they were collected.
Include the timeline plots from the poster.
Explain the different sequencing steps.

\section{Nextflow and nf-core}
Describe the nextflow workflow management system and the nf-core community.

\section{nf-core/circrna}
High-level overview

\subsection{circRNA detection}
What is the main task here?

\subsubsection{CIRIquant}
Describe CIRIquant

\subsubsection{CIRCexplorer2}
Describe CIRCexplorer2

\subsubsection{circRNA finder}
Describe circRNA finder

\subsubsection{DCC}
Describe DCC

\subsubsection{find circ}
Describe find circ

\subsubsection{MapSplice}
Describe MapSplice

\subsubsection{Segemehl}
Describe Segemehl

\subsection{circRNA annotation}
What is the main task here?

\subsubsection{GTF based annotation}
Describe GTF based annotation

\subsubsection{Database based annotation}
Describe database based annotation

\subsection{circRNA quantification}

Why is this important?

\subsubsection{psirc-quant}
Describe psirc-quant

\subsection{miRNA interaction analysis}
What can we learn from here?

\subsubsection{miRanda}
Describe miRanda

\subsubsection{TargetScan}
Describe TargetScan

\subsection{Downstream analyses}
\subsubsection{R-shiny}
\subsubsection{Dimensionality reduction}
\subsubsection{Pathway analysis}
\subsubsection{Differential expression analysis}
\subsubsection{Genome browser}
